---
title: "Untitled"
output: html_document
date: "2023-03-30"
editor_options: 
  chunk_output_type: console
---


This markdown file simply creates the price dataset merging the zoopla price

```{r}
library(tidyverse)
library(jsonlite)
library(readxl)

data_folder <- './data'
sales_prices <- file.path(data_folder, 'zoopla_extracts_sales_valid_property_id_expanded_2021')
rental_prices <- file.path(data_folder, 'zoopla_extracts_rent_valid_property_id_expanded_2021')
ONSPD_multi_files <- file.path(data_folder, "ONSPD/Data/multi_csv")
```


# Create a clean house price dataset

This block creates a cleaned house price dataset using the Zoopla property data from the UBDC. The data is subsetted to only include property types named in 'conv_prop_types'. The price data includes geographical information like postcode, LSOA, MSOA, Region code, etc, it also includes last listed price, property type and the Zoopla price ID.

A dataframe of summary statistics for each property type is also returned. 

```{r}
# Getting a list of directories within the `sales_prices` directory.
# Omitting the root directory from the list
price_data <- list.dirs(sales_prices, full.names = FALSE)[-1] %>% 
  map_df(~{
    # Printing the current directory
    print(.x)
    
    # Constructing the ONSPD file path based on the directory name
    ONSPD_file_path <- file.path(ONSPD_multi_files, paste0("ONSPD_NOV_2021_UK_", .x, ".csv"))
    
    # Reading the ONSPD data and selecting specific columns
    ONSPD_df <- read_csv(ONSPD_file_path, col_type = rep('c', 7), 
                         col_select = c('pcds', 'oslaua', 'osward','ctry', 'rgn', 'lsoa11', 'msoa11')) %>%
      # Replacing spaces in `pcds` column to create `pcu` column
      mutate(pcu = str_replace_all(pcds, " ", ""))
    
    # Getting a list of all CSV files in the current directory
    temp_df <- list.files(file.path(sales_prices, .x),  pattern = '.csv', full.names = TRUE) %>%
      # Reading and joining each CSV file with the ONSPD data
      map_df(~{
        read_csv(.x, col_types = c('d', 'c', 'd', 'c', 'c'), 
                col_select = c("property_id"  , 'property_type', 'price_last', 'pcu', 'category')) %>%
          left_join(ONSPD_df, by = join_by(pcu))
      })
  })

# Defining property types
all_prop_types<- c("Detached house", "Maisonette", "Terraced house", "Detached bungalow", 
"Semi-detached house", "Cottage", "Flat", "Bungalow", "Lodge", 
NA, "Land", "Farmhouse", "End terrace house", "Farm", "Equestrian property", 
"Country house", "Town house", "Barn conversion", "Terraced bungalow", 
"Villa", "Block of flats", "Mobile/park home", "Semi-detached bungalow", 
"Studio", "Link-detached house", "Light industrial", "Chalet", 
"Mews house", "Restaurant/cafe", "Retail premises", "Parking/garage", 
"Hotel/guest house", "Office", "Leisure/hospitality", "Houseboat", 
"Industrial", "Warehouse", "Finca", "Pub/bar", "Longère")

conv_prop_types <- c("Detached house", "Maisonette", "Terraced house", "Detached bungalow", 
"Semi-detached house", "Cottage", "Flat", "Bungalow", "Lodge", "Farmhouse", "End terrace house", 
"Country house", "Town house", "Barn conversion", "Terraced bungalow", 
"Villa",  "Semi-detached bungalow", 
"Studio", "Link-detached house",  "Chalet", 
"Mews house",  "Finca")

# Reading LAD names and selecting first two columns
LAD_names <- read_csv(file.path(data_folder, "ONSPD", "Documents", "LA_UA names and codes UK as at 04_21.csv")) %>%
  select(1:2)

# Filtering the price data for residential property types and joining with LAD names
price_data2 <- price_data %>%
  filter(category == "Residential", property_type %in% conv_prop_types) %>%
  left_join(LAD_names, by = c('oslaua' = 'LAD21CD'))

# Saving the filtered data as an RDS file
write_rds(price_data2, file.path(data_folder, "price_data.rds"))

# Calculating summary statistics for each property type in price_data
summary_statistics_property_types <- price_data %>%
  group_by(property_type) %>%
  summarise(
    mean = mean(price_last),
    median = median(price_last),
    sd = sd(price_last),
    max = max(price_last),
    min = min(price_last),
    counts = n()
  )

summary_statistics_property_types 

```

# Processing low-use property data and estimating average property prices


This code processes low-use property data from multiple sources, calculates summary statistics, and estimates the average property prices for different types of properties.


## loading data

```{r}
price_data2 <- read_rds(file.path(data_folder, "price_data.rds"))

low_use_file_paths <- list.files(file.path(data_folder, "low_use_LAD_files"), full.names = TRUE)

UK_pop <- read_xls(file.path(data_folder, 'population_data', 'ukpopestimatesmid2021on2021geographyfinal.xls'),
                   sheet = 7,
                   skip = 7)

```



## Process the low use data

```{r}
weighted_prices <- low_use_file_paths %>%
  map_df(~ {
    LAD_data <- read_csv(.x, col_types = c("c", "i", "f", "i", "d", "i", "f", "c", "c", "c", "f", "f", "c"))

    LAD_data <- LAD_data %>%
      group_by(MSOACD) %>%
      summarise(across(where(is.numeric), sum)) %>%
      left_join(LAD_data %>%
                 group_by(MSOACD) %>%
                 summarise(across(where(is.character), first)), by = "MSOACD") %>%
      select(-lsoa11cd)

    LAD_data %>%
      left_join(summarised_prices, by = c('MSOACD' = 'msoa11')) %>%
      mutate(low_use_ratio = low_use / all_properties)
  })

```

##  Create summary statistics for lsoa-level data

```{r}
lsoa_stats_LAD_level <- low_use_file_paths %>%
  map_df(~ {
    read_csv(.x, col_types = c("c", "i", "f", "i", "d", "i", "f", "c", "c", "c", "f", "f", "c")) %>%
      mutate(low_use_fract = low_use/sum(low_use),
             log_low_use_fract = log2(low_use_fract),
             log_low_use_fract = ifelse(is.finite(log_low_use_fract), log_low_use_fract, 0)) %>%
      summarise(Admin_district_code = first(Admin_district_code),
              max_low_use_ratio_lsoa = max(low_use_ratio),
              min_low_use_ratio_lsoa = min(low_use_ratio),
              entropy = -sum(low_use_fract * log_low_use_fract))
  })

```


```{r}
list.files(file.path(data_folder, "scotland"), full.names = TRUE)

LAD_data <- read_csv( "./data/scotland/Aberdeen City S12000033.csv"   , col_types = c("c", "i", "f", "i", "d", "i", "f", "c", "c", "c", "f", "f", "c")) %>%
  mutate(low_use_fract = low_use/sum(low_use),
         log_low_use_fract = log2(low_use_fract),
         log_low_use_fract = ifelse(is.finite(log_low_use_fract), log_low_use_fract, 0)) 

 LAD_data %>%
  mutate(low_use_fract = low_use/sum(low_use),
         log_low_use_fract = log2(low_use_fract),
         #The below line removes any infinite values caused by no low use property
         log_low_use_fract = ifelse(is.finite(log_low_use_fract), log_low_use_fract, 0)) %>%
  summarise(Admin_district_code = first(Admin_district_code),
    max_low_use_ratio_lsoa = max(low_use_ratio),
            min_low_use_ratio_lsoa = min(low_use_ratio),
            entropy = -sum(low_use_fract * log_low_use_fract))    


```



## summary statistics for prices

calculate the summary statistics including the population data.
The weighted dataset is saved at the end

```{r}
summarised_prices <- price_data2 %>%
  group_by(msoa11) %>%
  summarise(
    mean_price = mean(price_last),
    median_price = median(price_last),
    counts = n()
  )


weighted_price_LAD <- weighted_prices %>%
  group_by(Admin_district_code) %>%
  summarise(
    mean_price_low_use = weighted.mean(mean_price, low_use),
    mean_price_homes = weighted.mean(mean_price, homes),
    mean_price_all_properties = weighted.mean(mean_price, all_properties),
    low_use = sum(low_use),
    homes = sum(homes),
    all_properties = sum(all_properties),
    LADNM = first(LADNM),
    Country_code = first(Country_code),
    RGNNM = first(RGNNM)
  ) %>%
  select(Admin_district_code, LADNM, RGNNM, Country_code, low_use, homes, all_properties, 
         mean_price_low_use, mean_price_homes, mean_price_all_properties) %>%
  mutate(
    low_use_perc = low_use / all_properties,
    price_ratio = mean_price_low_use / mean_price_all_properties
  ) %>%
  left_join(lsoa_stats_LAD_level, by = 'Admin_district_code' ) %>%
  left_join(UK_pop %>%
  select('Admin_district_code' = Code,
         'pop' = `All ages`) , by = 'Admin_district_code')

# Save the processed data as a CSV file
write_csv(weighted_price_LAD, file.path(data_folder, 'weighted_price_LAD.csv'))
  
```



# old code

```{r}
price_data2 <- read_rds(file.path(data_folder, "price_data.rds"))


# Get a list of files in the 'low_use_LAD_files' directory
low_use_file_paths <-  c(
  list.files(file.path(data_folder, "low_use_LAD_files"), full.names = TRUE),
  list.files(file.path(data_folder, "low_use_LAD_files_bourne"), full.names = TRUE),
  list.files(file.path(data_folder, "scotland"), full.names = TRUE)
)

# Summarize prices by 'msoa11' and calculate mean, median, and counts
summarised_prices <- price_data2 %>%
  group_by(msoa11) %>%
  summarise(
    mean_price = mean(price_last),
    median_price = median(price_last),
    counts = n()
  )

# Obtain a list of files from three directories, and then process each file in that list
weighted_prices <- low_use_file_paths %>%
  map_df(~ {
    LAD_data <- read_csv(.x, col_types = c("c", "i", "f", "i", "d", "i", "f", "c", "c", "c", "f", "f", "c"))
    
    # Sum numeric columns and join with first character columns after grouping by MSOACD
    LAD_data <- LAD_data %>%
      group_by(MSOACD) %>%
      summarise(across(where(is.numeric), sum)) %>%
      left_join(LAD_data %>%
        group_by(MSOACD) %>%
        summarise(across(where(is.character), first)), by = "MSOACD") %>%
      select(-lsoa11cd)
    
    # Join with summarised prices and calculate a new ratio column
    LAD_data %>%
      left_join(summarised_prices, by = c('MSOACD' = 'msoa11')) %>%
      mutate(low_use_ratio = low_use / all_properties)
  })


#creates summary statistics from lsoa level data summarised to LAD level
#used to find high density areas of low use property
lsoa_stats_LAD_level <- low_use_file_paths %>%
  map_df(~ {
    read_csv(.x, 
             col_types = c("c", "i", "f", "i", "d", "i", "f", "c", "c", "c", "f", "f", "c")
             ) %>%
  mutate(low_use_fract = low_use/sum(low_use),
         log_low_use_fract = log2(low_use_fract),
         #The below line removes any infinite values caused by no low use property
         log_low_use_fract = ifelse(is.finite(log_low_use_fract), log_low_use_fract, 0)) %>%
  summarise(Admin_district_code = first(Admin_district_code),
    max_low_use_ratio_lsoa = max(low_use_ratio),
            min_low_use_ratio_lsoa = min(low_use_ratio),
            entropy = -sum(low_use_fract * log_low_use_fract))    

  })

#load the uk ppopulation data
UK_pop <- read_xls(file.path(data_folder, 'population_data','ukpopestimatesmid2021on2021geographyfinal.xls'), 
                  sheet = 7, skip = 7)

# Group by 'Admin_district_code' and calculate weighted means and other metrics
weighted_price_LAD <- weighted_prices %>%
  group_by(Admin_district_code) %>%
  summarise(
    mean_price_low_use = weighted.mean(mean_price, low_use),
    mean_price_homes = weighted.mean(mean_price, homes),
    mean_price_all_properties = weighted.mean(mean_price, all_properties),
    low_use = sum(low_use),
    homes = sum(homes),
    all_properties = sum(all_properties),
    LADNM = first(LADNM),
    Country_code = first(Country_code),
    RGNNM = first(RGNNM)
  ) %>%
  select(Admin_district_code, LADNM, RGNNM, Country_code, low_use, homes, all_properties, 
         mean_price_low_use, mean_price_homes, mean_price_all_properties) %>%
  mutate(
    low_use_perc = low_use / all_properties,
    price_ratio = mean_price_low_use / mean_price_all_properties
  ) %>%
  left_join(lsoa_stats_LAD_level, by = 'Admin_district_code' ) %>%
  left_join(UK_pop %>%
  select('Admin_district_code' = Code,
         'pop' = `All ages`) , by = 'Admin_district_code')

# Save the processed data as a CSV file
write_csv(weighted_price_LAD, file.path(data_folder, 'weighted_price_LAD.csv'))

```


```{r}
test <- weighted_price_LAD %>%
  filter(RGNNM=='Scotland')

ggplot(weighted_price_LAD %>% filter(entropy>2.5)
       , aes( x= entropy, y = low_use_perc)) + geom_point()

```



```{r}
# the fraction of population for each country
weighted_price_LAD %>% mutate('country'= stringr::str_sub(Admin_district_code, 1, 1)) %>%
  group_by(country) %>%
  summarise(pop = sum(pop))%>%
  mutate(total_pop = c(56536419,5479900,3105410),
         fract = pop/total_pop)

```

```{r}

ggplot(weighted_price_LAD, aes(x = low_use_perc, y = price_ratio)) + geom_point()

```


# Oldham Errors

A collection of a small number of errors
```{r}
# Set the value for .x, which seems to represent a file identifier or region
.x <- "OL"

# Print the identifier to the console
print(.x)

# Construct a file path to the corresponding ONSPD (Office for National Statistics Postcode Directory) CSV file
ONSPD_file_path <- file.path(ONSPD_multi_files, paste0("ONSPD_NOV_2021_UK_", .x, ".csv"))

# Read the ONSPD data from the constructed file path, select certain columns, and then mutate to get 'pcu' column
ONSPD_df <- read_csv(ONSPD_file_path, col_type = rep('c', 7), 
                     col_select = c('pcds', 'oslaua', 'osward', 
                                    'ctry', 'rgn', 'lsoa11', 'msoa11')) %>%
  mutate(pcu = str_replace_all(pcds, " ", ""))

# List CSV files within the corresponding directory in 'sales_prices'
# and then read each file, joining it with the previously read ONSPD data
temp_df <- list.files(file.path(sales_prices, .x), pattern = '.csv', full.names = TRUE) %>%
  map_df(~{
    # Read data from the CSV file, select certain columns, and then join with ONSPD data using 'pcu' as key
    read_csv(.x, col_types = c('d', 'c', 'd', 'c', 'c'), 
             col_select = c("property_id", 'property_type', 'price_last', 'pcu', 'category')) %>%
    left_join(ONSPD_df, by = "pcu")
  })

```


```{r}

```

