---
title: "Mapping"
output: html_document
date: "2023-01-16"
editor_options: 
  chunk_output_type: console
---


```{r}
library(tidyverse)
library(sf)
library(ggplot2)
library(stringr)
library(osmdata)
data_folder <- "./data"
```

# PLot all Lads xoloured by collection type

```{r}
all_lad_codes_blue_green <- list.files(file.path(data_folder, "low_use_LAD_files")) %>%
  str_extract(.,"[\\S]+$") %>% str_remove(., ".csv")

all_lad_codes_bourne <- list.files(file.path(data_folder, "low_use_LAD_files_bourne")) %>%
  str_extract(.,"[\\S]+$") %>% str_remove(., ".csv")

all_lad_codes_scotland <- list.files(file.path(data_folder, "scotland")) %>%
  str_extract(.,"[\\S]+$") %>% str_remove(., ".csv")

LAD_sf <- read_sf(file.path(data_folder, "LAD_shapefiles")) %>%
  mutate(type = case_when(
    LAD22CD %in% all_lad_codes_blue_green ~"BG towns",
    LAD22CD %in% all_lad_codes_bourne ~"Bourne 2019",
    str_sub(LAD22CD, 1, 1)=="S" ~ "ScotGov",
    TRUE ~ "No Data"
  )) %>%
  filter( str_sub(LAD22CD, 1, 1)!="N" )
```


```{r}
table(LAD_sf$type)


```

# Plot GB 

```{r}

weighted_price_LAD <- read_csv('./data/weighted_price_LAD.csv') %>%
  mutate(quartiles= cut(low_use_perc, 
                        breaks = quantile(low_use_perc, 
                        probs = seq(0, 1, 0.25)), 
                        labels = 1:4, 
                        include.lowest = TRUE))
  
  

ggplot() +
  geom_sf(data = LAD_sf, 
          mapping = aes(fill = type, linetype = NA)) + 
  coord_sf() +
  ggtitle("Data available for project by collection type")


LAD_sf %>% left_join(weighted_price_LAD %>%
    select(LAD22CD = Admin_district_code, quartiles)) %>%
  ggplot() +
  geom_sf(mapping = aes(fill = as.factor(quartiles), linetype = NA)) + 
  coord_sf() +
  scale_fill_viridis_d(na.value = "grey") +
  labs(title = 'LADs coloured by low-use Quartile, Q4 has the most low-use', fill= 'quartiles')

  
```


# PLot LSOA data

DO not try to plot the whole of the UK it would be horrible

```{r}
data_path = file.path(data_folder, "low_use_LAD_files_bourne", "Southwark E09000028.csv")

single_lad  = LSOA_sf %>% inner_join(read_csv(data_path), by = c("LSOA11CD" = "lsoa11cd" )) %>%
  mutate(perc = low_use_ratio * 100) %>%
  #convert to correct coordinate sytem
  st_transform(., 4326)

#get bounding box for LAD
lad_bbox <- st_bbox(single_lad)
  
```


```{r}
test <- LSOA_sf %>%filter(LADNM=='Southwark')

test %>% ggplot() + geom_sf()

library(leaflet)

pal <- colorNumeric(palette = "viridis", domain = single_lad$low_use)

m <- leaflet() %>%
  addTiles() %>%
  setView(lng = -0.07, lat = 51.47, zoom = 13)# %>%
  #addPolygons(data = single_lad, color = pal(single_lad$low_use), stroke = 1, opacity = 0.8)


m %>% addPolygons(
    label = ~low_use,  # the column you want to display
    labelOptions = labelOptions(
        style = list("background-color" = "white"),
        textOnly = TRUE,
        direction = 'auto'  # position of label can be 'auto', 'right', 'left', etc.
    ),
    highlightOptions = highlightOptions(
        weight = 5,
        color = "#666",
        fillOpacity = 0.7,
        bringToFront = TRUE
    )
)


leaflet() %>%
  addTiles() %>%
  setView(lng = -0.07, lat = 51.47, zoom = 13) %>%
  addPolygons(
    data = single_lad, 
    color = pal(single_lad$low_use), 
    stroke = 1, 
    opacity = 0.8,
    label = ~paste("LSOA Code:", LSOA11CD, " . ", "Low Use count:", low_use),
    labelOptions =labelOptions(
        style = list(
            "background-color" = "white",
            "font-size" = "14px"  # adjust the size as needed
        ),
        textOnly = TRUE,
        direction = 'auto'
    ),
    highlightOptions = highlightOptions(
        weight = 5,
        color = "#666",
        fillOpacity = 0.7,
        bringToFront = TRUE
    )
  )
```


```{r}


"East Riding of Yorkshire E06000011.csv"

LSOA_sf <- bind_rows(read_sf(file.path(data_folder, "LSOA_(Dec_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3")) %>%
  select(LSOA11CD),
  read_sf(file.path(data_folder, "SG_DataZoneBdry_2011")) %>%
  select(LSOA11CD = DataZone)
  ) %>%
  left_join(read_rds(file.path(data_folder, "lsoa_to_all_lookup.rds")), by = c("LSOA11CD"="lsoa11cd"))


file_path <- file.path(data_folder, "low_use_LAD_files_bourne/Cornwall E06000052.csv")


#If the shape file is merged onto the tibble it no longer counts as a geometric object.

#that is why the merging is done onto the shape file

single_lad  = LSOA_sf %>% inner_join(read_csv(file_path), by = c("LSOA11CD" = "lsoa11cd" )) %>%
  mutate(perc = low_use_ratio * 100)



```

#get deciles for all data set

I am going to colour the maps by absolute decile of low use percentage. to do this I need to load the whole dataset then find the deciles
```{r}
# Load all data into a list
all_data <- c('low_use_LAD_files', 'low_use_LAD_files_bourne', 'scotland') %>% 
  map(~{list.files(file.path(data_folder, .x), full.names = TRUE)}) %>%
  unlist %>%
  map(~ read_csv(.x)) %>%
  bind_rows()

# Calculate the combined perc
all_data <- all_data %>%
  mutate(
    low_use = ifelse(is.na(low_use), 0, low_use),
    low_use_ratio = ifelse(is.na(low_use_ratio), 0, low_use_ratio),
    perc = low_use_ratio * 100)

# Calculate deciles using the combined data
decile_breaks <- quantile(all_data$perc, probs = seq(0, 1, 0.1))


all_data <- all_data %>% mutate(
  decile = cut(perc, breaks = decile_breaks[0:10], labels = 1:9, include.lowest = TRUE)
  )



all_data %>%
  ggplot(aes(x = perc)) +geom_density( ) +
  labs(title=  "Percentage of LSOA low-use across all dataset", subtitle = "Mean value: 2.8, Median value: 1.9",
       x = "Percentage of LSOA that is low-use")

mean(all_data$perc)
median(all_data$perc)
```



```{r}

c('low_use_LAD_files', 'low_use_LAD_files_bourne', 'scotland') %>% 
  map(~{list.files(file.path(data_folder, .x), full.names = TRUE)}) %>%
  unlist

list.files(file.path(data_folder, 'low_use_LAD_files'))

lads_to_map <- c( file.path(data_folder,'low_use_LAD_files','West Devon E07000047.csv'),
                  file.path(data_folder,'low_use_LAD_files_bourne','West Suffolk E07000245.csv'),
                  file.path(data_folder,'low_use_LAD_files_bourne','Conwy W06000003.csv'),
                  file.path(data_folder,'scotland','East Lothian S12000010.csv') )

lads_to_map 

c('low_use_LAD_files', 'low_use_LAD_files_bourne', 'scotland') %>% 
  map(~{list.files(file.path(data_folder, .x), full.names = TRUE)}) %>%
  unlist %>% walk(~{
    
    temp <- read_csv(.x)
    single_lad  = LSOA_sf %>% 
      filter(Admin_district_code == unique(temp$Admin_district_code)[1]) %>%
      left_join(temp, by = c("LSOA11CD" = "lsoa11cd" )) %>%
      mutate(
        low_use = ifelse(is.na(low_use), 0, low_use),
        low_use_ratio = ifelse(is.na(low_use_ratio), 0, low_use_ratio),
        perc = low_use_ratio * 100,
        decile = cut(perc, breaks = decile_breaks[0:10], labels = 1:9, include.lowest = TRUE),
        decile = ifelse(is.na(decile), 10, decile),
        decile = factor(decile, levels = 1:10))
    
    LADNM <- basename(.x) %>% str_remove(., "\\s\\w\\d.*")
    plot_title <- paste("Low use properties as a percent of LSOA in", LADNM)
    
    ggplot() +
      geom_sf(data = single_lad, mapping = aes(fill = decile)) + 
      coord_sf() +
      scale_fill_viridis_d(name = "Decile", 
                           breaks = as.character(1:10), 
                           labels = 1:10.,
                           drop = FALSE) +
      ggtitle(plot_title)
    
    
    ggsave(file.path(data_folder, 'maps', paste(basename(.x) %>% str_replace(., ".csv", ".pdf"))))
    
  })
```



```{r}
temp <- read_csv("./data/low_use_LAD_files/Gwynedd W06000002.csv" )
  single_lad  = LSOA_sf %>% 
    filter(Admin_district_code == unique(temp$Admin_district_code)[1]) %>%
    left_join(temp, by = c("LSOA11CD" = "lsoa11cd" )) %>%
  mutate(
    low_use = ifelse(is.na(low_use), 0, low_use),
    low_use_ratio = ifelse(is.na(low_use_ratio), 0, low_use_ratio),
    perc = low_use_ratio * 100,
    decile = cut(perc, breaks = decile_breaks[0:10], labels = 1:9, include.lowest = TRUE),
     decile = ifelse(is.na(decile), 10, decile),
    decile = factor(decile, levels = 1:10))

LADNM <- basename('kkkk') %>% str_remove(., "\\s\\w\\d.*")
plot_title <- paste("Low use properties as a percent of LSOA in", LADNM)

ggplot() +
    geom_sf(data = single_lad, mapping = aes(fill = decile)) + 
    coord_sf() +
    scale_fill_viridis_d(name = "Decile", 
                         breaks = as.character(1:10), 
                         labels = 1:10.,
                         drop = FALSE) +
    ggtitle(plot_title)

```


South Devon, East Suffolk, Conwy, East Lothian

# map with roads

```{r}



low_use_map_plot <- function(data_path, LSOA_sf,  title = "plot title"){
  
  ##
  ## Create a plot of the named lad using the file path
  ## depends on ggplot2, sf, osmdata, tidyverse
  ##
  
  #If the shape file is merged onto the tibble it no longer counts as a geometric object.
#that is why the merging is done onto the shape file
single_lad  = LSOA_sf %>% inner_join(read_csv(data_path), by = c("LSOA11CD" = "lsoa11cd" )) %>%
  mutate(perc = low_use_ratio * 100) %>%
  #convert to correct coordinate sytem
  st_transform(., 4326)

#get bounding box for LAD
lad_bbox <- st_bbox(single_lad)

LAD_major <- lad_bbox %>%
  opq(timeout = 50) %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", "secondary")) %>%
  osmdata_sf()

LAD_minor <- lad_bbox %>%
  opq(timeout = 50) %>%
  add_osm_feature(key = "highway", value = c("tertiary", "residential")) %>%
  osmdata_sf()

out <- ggplot()+
      geom_sf(data = single_lad, mapping = aes(fill = perc,  linetype = NA)) + 
  coord_sf() +
  geom_sf(data = LAD_major$osm_lines,
          inherit.aes = FALSE,
          color = "orange",
          size = 0.2) +
  geom_sf(data = LAD_minor$osm_lines,
          inherit.aes = FALSE,
          color = "orange",
          size = 0.2) +
  scale_fill_viridis_b() +
  labs(title = title, fill = "Percent")

return(out)
  
}

```



```{r}
g <- low_use_map_plot(data_path = file.path(data_folder, "low_use_LAD_files", 
                                            "Gwynedd W06000002.csv"), 
                 LSOA_sf,  title = "Low use properties as percent of LSOA in Tower Hamlets" )

g


g2 <- low_use_map_plot(data_path = file.path(data_folder, "scotland", "Argyll and Bute S12000035.csv"), 
                 LSOA_sf,  title = "Low use properties as percent of MSOA in East Riding" )

g2

```



# Create Hex binned uk map

The issue of certain local authorities dominating or being swamped by others due to physical size can be overcome using hexbins




```{r}

library(jsonlite)

data <- read_json(file.path(data_folder, 'hex_mapping' ,'uk-local-authority-districts-2023.hexjson'))
hexes <- data$hexes

df <- enframe(data$hexes, name = "id") 
# Spread the nested list into multiple columns
df <- df %>% unnest_wider(value) %>%
  filter(region != "N92000002")  %>%
  left_join(weighted_price_LAD %>%
    select(Admin_district_code, quartiles), by = c('id' = 'Admin_district_code')) %>%
  mutate(type = case_when(
    id %in% all_lad_codes_blue_green ~"BG towns",
    id %in% all_lad_codes_bourne ~"Bourne 2019",
    str_sub(id, 1, 1)=="S" ~ "ScotGov",
    TRUE ~ "No Data"
  ),
  quartiles = as.factor(region))


hex_vertices <- function(q, r, layout="odd-r") {
  size = 1 # The size has no impact on the plot but is included in case adding it 
  #as an argument later is useful
  q_adjusted = q
  r_adjusted = r
  
  # Validate layout input
  valid_layouts <- c("odd-r", "even-r", "odd-q", "even-q")
  if (!layout %in% valid_layouts) {
    stop("Invalid layout specified. Choose from: ", paste(valid_layouts, collapse=", "))
  }
  
  # Adjust for the layout
  if (layout == "odd-r" && (r %% 2) == 1) {
    q_adjusted = q + 0.5
  } else if (layout == "even-r" && (r %% 2) == 0) {
    q_adjusted = q + 0.5
  } else if (layout == "odd-q" && (q %% 2) == 1) {
    r_adjusted = r + 0.5
  } else if (layout == "even-q" && (q %% 2) == 0) {
    r_adjusted = r + 0.5
  }
  
  if(layout %in% c("odd-r", "even-r")) { # Pointy-topped hex grid
    # Convert (q_adjusted, r_adjusted) directly to Cartesian x, y
    x = size * sqrt(3) * q_adjusted
    y = size * 3/2 * r_adjusted
    # Calculate vertices for pointy-topped hex
    angles <- seq(0, 5*pi/3, by = pi/3)
    data.frame(x = x + size * sin(angles), 
               y = y + size * cos(angles))
  } else { # Flat-topped hex grid
    # Convert (q_adjusted, r_adjusted) directly to Cartesian x, y
    x = size * 3/2 * q_adjusted
    y = size * sqrt(3) * r_adjusted
    # Calculate vertices for flat-topped hex
    angles <- seq(pi/6, 13*pi/6, by = pi/3)
    data.frame(x = x + size * cos(angles), 
               y = y + size * sin(angles))
  }
}



df_vertices <- df %>%
  rowwise() %>%
  do(hex_vertices(.$q, .$r,  'even-r'))  %>% ungroup %>%
  mutate(id = rep(df$id, each=6), 
         fill_value = rep(df$quartiles, each=6))

g <- ggplot(df_vertices, aes(x=x, y=y, group=id, fill=fill_value )) +
  geom_polygon( color="black") +
  theme_minimal() +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
    axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


g+  labs(title = 'LADs coloured by low-use Quartile, Q4 has the most low-use', fill = 'Quartiles',
       x = NULL, y = NULL)  
+
  scale_fill_viridis_d(na.value = "grey") 


g
```

